Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses

### 2024-12-13 manuscript_agu.tex
Fixing grammar

\begin{abstract}
Global storm-resolving models (GSRMs) are the upcoming global climate models. One of them is a 5-km Icosahedral Nonhydrostatic Weather and Climate Model (ICON). Its high resolution means that parameterizations of convection and clouds, including subgrid-scale clouds, are omitted, relying on explicit simulation but still utilizing microphysics and turbulence parameterizations. Standard-resolution (10--100~km) models, which use convection and cloud parameterizations, have substantial cloud biases over the Southern Ocean (SO), adversely affecting radiation and sea surface temperature. The SO is dominated by low clouds, which cannot be observed accurately from space due to overlapping clouds, attenuation, and ground clutter. We evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses using about 2400 days of lidar observations and 2300 radiosonde profiles from 31 voyages and Macquarie Island station during 2010--2021, compared with the models using a ground-based lidar simulator. We found that ICON and the reanalyses underestimate the total cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate the cloud occurrence peak at about 500 m, potentially explained by their lifting condensation levels being too high. The reanalyses strongly underestimate fog or near-surface clouds, and MERRA-2 underestimates cloud occurrence at almost all heights. Outgoing shortwave radiation is overestimated in the reanalyses, implying a ``too few, too bright'' cloud problem. Thermodynamic conditions are relatively well represented, but ICON is less stable than observations and MERRA-2 is too humid. SO cloud biases are a substantial issue in the GSRM, but it matches the observations better than the lower-resolution reanalyses.
\end{abstract}

@@

The Southern Ocean (SO) is known to be a problematic region for climate model biases \cite{schuddeboom2021,hyder2018,cesana2022,zhao2022} due to a lack of surface and in situ observations and being a lower priority region for numerical weather prediction (NWP) and climate model development because of its distance from populated areas. Nevertheless, radiation biases and changes over an area of its size have a substantial influence on the global climate \cite{rintoul2011}, such as affecting the Earth radiation balance, ocean heat, and carbon uptake \cite{williams2023}, and the SO is also an important part of the global ocean conveyor belt \cite{wang2014b}. In general, marine clouds have a disproportionate effect on top of atmosphere (TOA) SW radiation due to the relatively low albedo of the sea surface. The relative longitudinal symmetry of the SO means that model cloud biases tend to be similar across longitudes.

@@

The main aim of this study is to evaluate the GSRM version of ICON. ICON is developed and maintained jointly by Deutscher Wetterdienst, Max-Planck-Institute for Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of Technology, and the Center for Climate Systems Modeling. Previous studies have identified substantial large-scale biases in climate model clouds over the SO, affecting sea surface temperature and the Earth’s albedo. Our aim is to quantify how well the GSRM ICON simulates clouds in this region, particularly in light of the fact that subgrid-scale clouds and convection are not parameterized in this model. This region is mostly dominated by boundary layer clouds generated by shallow convection, and these are problematic to observe by spaceborne lidars and radars, which are affected by attenuation by overlapping and thick clouds \cite{mace2009,medeiros2010} and ground clutter \cite{marchand2008}, respectively. Specifically, the radar on CloudSat and lidar on CALIPSO (neither of which are now operational) are affected by the above mentioned issues, resulting in a strong underestimation of cloud occurrence below 2 km relative to ground-based lidar observations \cite{mcerlich2021}. We hypothesize that this, in turn, can lead to systematic biases in low clouds in climate models, which are frequently evaluated against CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as these are usually parameterized in their atmospheric component and also in regions where input observations are sparse. This makes them a problematic reference for clouds over the SO, and any biases relative to a reanalysis should be interpreted with caution. Instead, we chose to use a large set of ship-based observations conducted with ceilometers and lidars on board the RV \emph{Polarstern} and other voyages and stations as a reference for the model evaluation.

@@

The campaigns contained ceilometer observations captured by the Vaisala CL51 and CT25K, and the Lufft CHM~15k, described in detail below (Sections~\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power, near-infrared, vertically pointing lidar principally designed to measure cloud base, but they also measure the full vertical structure of clouds as long as the laser signal is not attenuated by thick clouds, which can be used to infer additional information such as a cloud mask and cloud occurrence by height. We note that during the MICRE campaign, the ceilometers Vaisala CT25K and CL51 were installed at the Macquarie Island station concurrently, but in our analysis we only used the CT25K data obtained from the Atmospheric Radiation Measurement (ARM) data archive.

@@

Fig.~\ref{fig:cyclone-stability}a, b show the geographical distribution of the fraction of cyclonic days as determined by the cyclone tracking algorithm applied to the ERA5 reanalysis and ICON data (Section~\ref{sec:cyclone-stability}). As expected, the strongest cyclonic activity is in the high-latitude SO zone, and is relatively zonally symmetric at all latitudes. The pattern matches reasonably well with \citeA{hoskins2005}. While both reanalysis and the model agree within about 8\% in most areas, ICON is prevailingly more cyclonic by about 4\%. There are clear differences, particularly in the highest occurrence rate regions, such as around Cape Adare, which is up to 20\% more cyclonic in ICON, and the Weddell and Bellingshausen Seas, where ICON is less cyclonic by up to 10\%. These differences might, however, stem from the relatively short time periods of comparison (4 years) and the fact that the model is free-running.

@@

For all observations considered (Fig.~\ref{fig:cloud-occurrence}a), the data show cloud occurrence peaking nearly at the surface, whereas the models show a higher peak (at about 500~m). The models generally underestimate the total cloud fraction by 10--20\% and show a strong reduction in cloud occurrence near the surface, which is not identified in the observations. ICON and ERA5 overestimate cloud occurrence at their peak (between 0--1~km). Above 1~km, ICON and MERRA-2 underestimate cloud occurrence, but ERA5 is accurate to about 3\% or less. The exaggerated peak in models is partly explained by the lifting condensation level (LCL) distribution, which peaks about 200~m higher in the models than in the observations (nearly at the surface), although this is not very pronounced. This is indicative of near-surface relative humidity being often close to saturation in the observations but not in the models.

@@

In Fig.~\ref{fig:cloud-occurrence}, we also display the mean outgoing shortwave and longwave top of atmosphere radiation, whose calculation is described in Section~\ref{sec:ceres}. In observations, these come from daily mean CERES measurements averaged over the voyage tracks or a station location, whereas in the models they come from daily means of TOA radiation in the model output averaged over the same location and time periods. In the free-running ICON model, the time period is mapped onto the available years, as explained in Section~\ref{sec:icon}.

In the general case (Fig.~\ref{fig:cloud-occurrence}a), ICON underestimates the outgoing SW radiation by 26 Wm$^{-2}$ and the MERRA-2 and ERA5 reanalyses overestimate it by 6 and 14 Wm$^{-2}$, respectively. While in ICON, this is in line with the underestimated total cloud fraction of 10\%, in the reanalyses this is the opposite result to that expected from the underestimated total cloud fraction of about 20\%. The likely explanation is an overestimated cloud albedo, compensating for the lack of cloud area.

We note that the the radiative transfer calculations used in the lidar simulator mean that the impact of both cloud phase and cloud fraction are convolved to produce the cloud mask. Therefore, the cloud occurrence is not affected by any cloud phase biases as long as the cloud is optically thick enough to be detected and the laser signal is not too attenuated. However, a combination of underestimated total cloud fraction and overestimated outgoing SW at TOA is indicative of an overestimated cloud albedo due to either cloud liquid and ice water content, cloud phase, droplet or ice crystal size distribution, shape or orientation of ice crystals, or cloud overlap.

In contrast, LW radiation has much smaller biases than SW radiation, which is expected due to the prevailing low-level clouds having similar temperature as the surface. In ICON, the outgoing LW radiation is overestimated by 8\%, which could be caused by an underestimated total cloud fraction exposing a larger sea surface area to cooling to space, which is typically warmer than the atmospheric temperature at 0--2~km, where most of the clouds are located. In the MERRA-2 and ERA5 reanalyses, the LW biases are also slightly positive 4 and 7~Wm$^{-2}$, respectively. This is again in line with the underestimated total cloud fraction by about 20\%. However, if the clouds are too thick, as expected from the SW results, this might also provide a compensating effect, in which too small a cloud area is counteracted by greater thermal emissivity, thus reducing the outgoing LW radiation more relative to thinner clouds. For thin clouds, the outgoing TOA LW radiation originates both from the warmer surface (partly blocked by the clouds) and the clouds, whereas for thick clouds, the outgoing TOA LW radiation originates mostly from the colder-than-surface clouds.

In all the subsets (Fig.~\ref{fig:cloud-occurrence}b--g), the same type biases are observed, namely the outgoing SW radiation is underestimated in ICON and overestimated in MERRA-2 and ERA5, and the outgoing LW radiation is overestimated in all the models. Even though the total cloud fraction is lower by 7\% over the high-latitude SO than the low-latitude SO, the outgoing SW radiation is much greater by 41~Wm$^{-2}$, implying a much greater cloud albedo over the high-latitude SO. ICON model output displays the same contrast between these two regions in the total cloud fraction and SW radiation, but the outgoing SW radiation difference between the regions is much smaller (16~Wm$^{-2}$). The reanalyses do not show this type of contrast between the regions. The physical reason for this might be that the prevalence of fog or low-level clouds over the low-latitude SO and their relative lack over the high-latitude SO in observations is not reproduced in the models (Fig.~\ref{fig:cloud-occurrence}b--c).

@@

We also analyzed the daily cloud cover (total cloud fraction) distribution. This is a measure of cloudiness, irrespective of height, calculated over the course of a day (UTC). A cloud detected at any height means that the lidar profile was classified as cloudy; otherwise, it was classified as clear sky. When all profiles in a day are taken together, the cloud cover for the day is defined as the fraction of cloudy profiles in the total number of profiles, expressed in oktas (multiples of 1/8). The same calculation is done for the lidar observations as for the simulated lidar profiles. We use the term ``okta'' independently of its use in instantaneous synoptic observations, and here it simply means 1/8 (0.125\%) of the daily cloud cover.

@@

The time period of ICON is relatively short, with only four full years of simulation available. Moreover, the simulation is free-running and ocean-coupled, which means that observations had to be temporally mapped to this time period (at the same time relative to the start of the year) for the comparison. For these reasons, one can expect the results to be slightly different due to reasons unrelated to model biases, such as different weather conditions, partially accounted for by the cyclone and stability subsetting, and the phase of climate oscillations such as the ENSO in the observations and the model. The interannual variability in cloud occurrence in ICON can be seen in Fig.~\ref{fig:cloud-occurrence-panel}, where each year in ICON is represented by a separate line. The interannual variability tends to be substantially smaller than the biases, and thus is unlikely to have a strong impact on the main findings.

@@

We have attempted to remove lidar profiles with precipitation, which could not be properly simulated with the lidar simulator (Section~\ref{sec:ann}). However, the approach was limited by the relatively low sensitivity of the ANN (65\%) and the fact that we had to choose a fixed threshold for surface precipitation flux in the model and reanalyses, which might not correspond to detection by the ANN applied to observations. We also made no attempt to remove profiles with precipitation which did not reach the surface. The above reasons may result in an artificial bias in the comparison, though we expect this to be much smaller than the identified model biases.

@@

Our main finding corroborates previous findings of large boundary layer cloud biases in models and their subsequent effect on the radiative transfer. For example, low- and mid-level clouds in the cold-air sector of cyclones were identified as being responsible for most of the SW bias in \citeA{bodas-salcedo2012}. This understanding was refined in \citeA{bodas-salcedo2014}, which highlighted that the SW bias was associated with an incorrectly simulated mid-level cloud regime, which occurred in regions where cloud with tops at mid-level and low-levels occurred. Our results align less well with more recent work by \citeA{ramadoss2024} which shows persistent shortwave radiative bias over the Southern Ocean are associated with incorrect cloud phase representation. While \citeA{fiddes2024} suggest biases in the liquid water path are the largest contributor to the cloud radiative bias over the Southern Ocean. Our general finding applies to the new GSRM ICON, but the biases are generally lower than in the reanalyses, despite the reanalyses having the advantage of assimilation of the observed meteorological conditions. The GSRM has, on the other hand, the advantage of a much higher spatial resolution and, to a limited extent, explicit calculation of traditionally subgrid-scale processes such as convection.

@@

The radiosonde observations indicate that the LCL is too high in ICON and reanalyses, which is probably responsible for the higher peak of clouds in the models and the lack of near-surface clouds or fog. The radiosonde comparison, however, does not seem to explain cloud biases at higher altitudes, which is perhaps suggestive of biases in the influence of the liquid water path in the models relative to the reality. MERRA-2 is too moist at all heights. ICON also exhibits smaller internal variability than the radiosonde observations. Overall, the radiosonde comparison only partially explains the identified cloud biases, and other physical causes are likely contributing. This warrants further investigation, especially of ocean--atmosphere fluxes, shallow convection, and boundary layer turbulence. The lack of parameterized subgrid-scale convection in ICON could be a substantial issue even at the 5-km resolution.

>>>>

\begin{abstract}
Global storm-resolving models (GSRMs) are the upcoming global climate models. One of them is a 5-km Icosahedral Nonhydrostatic Weather and Climate Model (ICON). Its high resolution means that parameterizations of convection and clouds, including subgrid-scale clouds, are omitted, relying on explicit simulation but still utilizing microphysics and turbulence parameterizations. Standard-resolution (10--100~km) models, which use convection and cloud parameterizations, have substantial cloud biases over the Southern Ocean (SO), adversely affecting radiation and sea surface temperature. The SO is dominated by low clouds, which cannot be observed accurately from space due to overlapping clouds, attenuation, and ground clutter. We evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses using about 2400 days of lidar observations and 2300 radiosonde profiles from 31 voyages and Macquarie Island station during 2010--2021, compared with the models using a ground-based lidar simulator. We found that ICON and the reanalyses underestimate the total cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate the cloud occurrence peak at about 500 m, potentially explained by their lifting condensation levels being too high. The reanalyses strongly underestimate fog or near-surface clouds, and MERRA-2 underestimates cloud occurrence at almost all heights. Outgoing shortwave radiation is overestimated in the reanalyses, implying a ``too few, too bright'' cloud problem. Thermodynamic conditions are relatively well represented, but ICON is less stable than observations, and MERRA-2 is too humid. SO cloud biases are a substantial issue in the GSRM, but it matches the observations better than the lower-resolution reanalyses.
\end{abstract}

@@

The Southern Ocean (SO) is known to be a problematic region for climate model biases \cite{schuddeboom2021,hyder2018,cesana2022,zhao2022} due to a lack of surface and in situ observations and being a lower priority region for numerical weather prediction (NWP) and climate model development because of its distance from populated areas. Nevertheless, radiation biases and changes over an area of its size have a substantial influence on the global climate \cite{rintoul2011}, such as affecting the Earth radiation balance, ocean heat, and carbon uptake \cite{williams2023}, and the SO is also an important part of the global ocean conveyor belt \cite{wang2014b}. In general, marine clouds have a disproportionate effect on top-of-atmosphere (TOA) SW radiation due to the relatively low albedo of the sea surface. The relative longitudinal symmetry of the SO means that model cloud biases tend to be similar across longitudes.

@@

The main aim of this study is to evaluate the GSRM version of ICON. ICON is developed and maintained jointly by Deutscher Wetterdienst, Max-Planck-Institute for Meteorology, Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute of Technology, and the Center for Climate Systems Modeling. Previous studies have identified substantial large-scale biases in climate model clouds over the SO, affecting sea surface temperature and the Earth’s albedo. Our aim is to quantify how well the GSRM ICON simulates clouds in this region, particularly in light of the fact that subgrid-scale clouds and convection are not parameterized in this model. This region is mostly dominated by boundary layer clouds generated by shallow convection, and these are problematic to observe by spaceborne lidars and radars, which are affected by attenuation by overlapping and thick clouds \cite{mace2009,medeiros2010} and ground clutter \cite{marchand2008}, respectively. Specifically, the radar on CloudSat and lidar on CALIPSO (neither of which are now operational) are affected by the above-mentioned issues, resulting in a strong underestimation of cloud occurrence below 2 km relative to ground-based lidar observations \cite{mcerlich2021}. We hypothesize that this, in turn, can lead to systematic biases in low clouds in climate models, which are frequently evaluated against CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as these are usually parameterized in their atmospheric component and also in regions where input observations are sparse. This makes them a problematic reference for clouds over the SO, and any biases relative to a reanalysis should be interpreted with caution. Instead, we chose to use a large set of ship-based observations conducted with ceilometers and lidars on board the RV \emph{Polarstern} and other voyages and stations as a reference for the model evaluation.

@@

The campaigns contained ceilometer observations captured by the Vaisala CL51, CT25K, and the Lufft CHM~15k, described in detail below (Sections~\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power, near-infrared, vertically pointing lidar principally designed to measure cloud base, but they also measure the full vertical structure of clouds as long as the laser signal is not attenuated by thick clouds, which can be used to infer additional information such as a cloud mask and cloud occurrence by height. We note that during the MICRE campaign, the ceilometers Vaisala CT25K and CL51 were installed at the Macquarie Island station concurrently, but in our analysis we only used the CT25K data obtained from the Atmospheric Radiation Measurement (ARM) data archive.

@@

Fig.~\ref{fig:cyclone-stability}a, b show the geographical distribution of the fraction of cyclonic days as determined by the cyclone tracking algorithm applied to the ERA5 reanalysis and ICON data (Section~\ref{sec:cyclone-stability}). As expected, the strongest cyclonic activity is in the high-latitude SO zone and is relatively zonally symmetric at all latitudes. The pattern matches reasonably well with \citeA{hoskins2005}. While both reanalysis and the model agree within about 8\% in most areas, ICON is prevailingly more cyclonic by about 4\%. There are clear differences, particularly in the highest occurrence rate regions, such as around Cape Adare, which is up to 20\% more cyclonic in ICON, and the Weddell and Bellingshausen Seas, where ICON is less cyclonic by up to 10\%. These differences might, however, stem from the relatively short time periods of comparison (4 years) and the fact that the model is free-running.

@@

For all observations considered (Fig.~\ref{fig:cloud-occurrence}a), the data show cloud occurrence peaking nearly at the surface, whereas the models show a higher peak (at about 500~m). The models generally underestimate the total cloud fraction by 10--20\% and show a strong reduction in cloud occurrence near the surface, which is not identified in the observations. ICON and ERA5 overestimate cloud occurrence at their peak (between 0 and 1~km). Above 1~km, ICON and MERRA-2 underestimate cloud occurrence, but ERA5 is accurate to about 3\% or less. The exaggerated peak in models is partly explained by the lifting condensation level (LCL) distribution, which peaks about 200~m higher in the models than in the observations (nearly at the surface), although this is not very pronounced. This is indicative of near-surface relative humidity being often close to saturation in the observations but not in the models.

@@

In Fig.~\ref{fig:cloud-occurrence}, we also display the mean outgoing shortwave and longwave top-of-atmosphere radiation, whose calculation is described in Section~\ref{sec:ceres}. In observations, these come from daily mean CERES measurements averaged over the voyage tracks or a station location, whereas in the models they come from daily means of TOA radiation in the model output averaged over the same location and time periods. In the free-running ICON model, the time period is mapped onto the available years, as explained in Section~\ref{sec:icon}.

In the general case (Fig.~\ref{fig:cloud-occurrence}a), ICON underestimates the outgoing SW radiation by 26 Wm$^{-2}$, and the MERRA-2 and ERA5 reanalyses overestimate it by 6 and 14 Wm$^{-2}$, respectively. While in ICON, this is in line with the underestimated total cloud fraction of 10\%, in the reanalyses this is the opposite result to that expected from the underestimated total cloud fraction of about 20\%. The likely explanation is an overestimated cloud albedo, compensating for the lack of cloud area.

We note that the radiative transfer calculations used in the lidar simulator mean that the impact of both cloud phase and cloud fraction are convolved to produce the cloud mask. Therefore, the cloud occurrence is not affected by any cloud phase biases as long as the cloud is optically thick enough to be detected, and the laser signal is not too attenuated. However, a combination of underestimated total cloud fraction and overestimated outgoing SW at TOA is indicative of an overestimated cloud albedo due to either cloud liquid and ice water content, cloud phase, droplet or ice crystal size distribution, shape or orientation of ice crystals, or cloud overlap.

In contrast, LW radiation has much smaller biases than SW radiation, which is expected due to the prevailing low-level clouds having similar temperature as the surface. In ICON, the outgoing LW radiation is overestimated by 8\%, which could be caused by an underestimated total cloud fraction exposing a larger sea surface area to cooling to space, which is typically warmer than the atmospheric temperature at 0--2~km, where most of the clouds are located. In the MERRA-2 and ERA5 reanalyses, the LW biases are also slightly positive, 4 and 7~Wm$^{-2}$, respectively. This is again in line with the underestimated total cloud fraction by about 20\%. However, if the clouds are too thick, as expected from the SW results, this might also provide a compensating effect, in which too small a cloud area is counteracted by greater thermal emissivity, thus reducing the outgoing LW radiation more relative to thinner clouds. For thin clouds, the outgoing TOA LW radiation originates both from the warmer surface (partly blocked by the clouds) and the clouds, whereas for thick clouds, the outgoing TOA LW radiation originates mostly from the colder-than-surface clouds.

In all the subsets (Fig.~\ref{fig:cloud-occurrence}b--g), the same type of biases are observed, namely the outgoing SW radiation is underestimated in ICON and overestimated in MERRA-2 and ERA5, and the outgoing LW radiation is overestimated in all the models. Even though the total cloud fraction is lower by 7\% over the high-latitude SO than the low-latitude SO, the outgoing SW radiation is much greater by 41~Wm$^{-2}$, implying a much greater cloud albedo over the high-latitude SO. The ICON model output displays the same contrast between these two regions in the total cloud fraction and SW radiation, but the outgoing SW radiation difference between the regions is much smaller (16~Wm$^{-2}$). The reanalyses do not show this type of contrast between the regions. The physical reason for this might be that the prevalence of fog or low-level clouds over the low-latitude SO and their relative lack over the high-latitude SO in observations is not reproduced in the models (Fig.~\ref{fig:cloud-occurrence}b--c).

@@

We also analyzed the daily cloud cover (total cloud fraction) distribution. This is a measure of cloudiness, irrespective of height, calculated over the course of a day (UTC). A cloud detected at any height means that the lidar profile was classified as cloudy; otherwise, it was classified as a clear sky. When all profiles in a day are taken together, the cloud cover for the day is defined as the fraction of cloudy profiles in the total number of profiles, expressed in oktas (multiples of 1/8). The same calculation is done for the lidar observations as for the simulated lidar profiles. We use the term ``okta'' independently of its use in instantaneous synoptic observations, and here it simply means 1/8 (0.125\%) of the daily cloud cover.

@@

The time period of ICON is relatively short, with only four full years of simulation available. Moreover, the simulation is free-running and ocean-coupled, which means that observations had to be temporally mapped to this time period (at the same time relative to the start of the year) for the comparison. For these reasons, one can expect the results to be slightly different due to reasons unrelated to model biases, such as different weather conditions, partially accounted for by the cyclone and stability subsetting, and the phase of climate oscillations such as the ENSO in the observations and the model. The interannual variability in cloud occurrence in ICON can be seen in Fig.~\ref{fig:cloud-occurrence-panel}, where each year in ICON is represented by a separate line. The interannual variability tends to be substantially smaller than the biases and thus is unlikely to have a strong impact on the main findings.

@@

We have attempted to remove lidar profiles with precipitation, which could not be properly simulated with the lidar simulator (Section~\ref{sec:ann}). However, the approach was limited by the relatively low sensitivity of the ANN (65\%) and the fact that we had to choose a fixed threshold for surface precipitation flux in the model and reanalyses, which might not correspond to detection by the ANN applied to observations. We also made no attempt to remove profiles with precipitation that did not reach the surface. The above reasons may result in an artificial bias in the comparison, though we expect this to be much smaller than the identified model biases.

@@

Our main finding corroborates previous findings of large boundary layer cloud biases in models and their subsequent effect on the radiative transfer. For example, low- and mid-level clouds in the cold-air sector of cyclones were identified as being responsible for most of the SW bias in \citeA{bodas-salcedo2012}. This understanding was refined in \citeA{bodas-salcedo2014}, which highlighted that the SW bias was associated with an incorrectly simulated mid-level cloud regime, which occurred in regions where clouds with tops at mid-level and low-levels occurred. Our results align less well with more recent work by \citeA{ramadoss2024}, which shows persistent shortwave radiative biases over the Southern Ocean are associated with incorrect cloud phase representation. While \citeA{fiddes2024} suggest biases in the liquid water path are the largest contributor to the cloud radiative bias over the Southern Ocean. Our general finding applies to the new GSRM ICON, but the biases are generally lower than in the reanalyses, despite the reanalyses having the advantage of assimilation of the observed meteorological conditions. The GSRM has, on the other hand, the advantage of a much higher spatial resolution and, to a limited extent, explicit calculation of traditionally subgrid-scale processes such as convection.

@@

The radiosonde observations indicate that the LCL is too high in ICON and reanalyses, which is probably responsible for the higher peak of clouds in the models and the lack of near-surface clouds or fog. The radiosonde comparison, however, does not seem to explain cloud biases at higher altitudes, which is perhaps suggestive of biases in the influence of the liquid water path in the models relative to reality. MERRA-2 is too moist at all heights. ICON also exhibits smaller internal variability than the radiosonde observations. Overall, the radiosonde comparison only partially explains the identified cloud biases, and other physical causes are likely contributing. This warrants further investigation, especially of ocean--atmosphere fluxes, shallow convection, and boundary layer turbulence. The lack of parameterized subgrid-scale convection in ICON could be a substantial issue even at the 5-km resolution.
